---
title: "DebugOverlap"
output: html_document
date: "2025-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(pgenlibr)
```

## Simple R Example 

```{r}
L <- 5000
M <- 100

# Random r 
r <- runif(L)
  
# Random "Genotypes"
G <- matrix(NA, nrow = M, ncol = L)
for (l in 1:L) {
  G[,l] <- rnorm(M, 0, 1)
}
  
# Scale vectors
r <- scale(r)
G <- scale(G)
```


Definition of $H$

$$H = \frac{\mathbb{E}_{M} \left[Cov(r, G_{i \cdot})^2 \right] } {\mathbb{E}_{M} \left[Var(G_{i \cdot}) \right]}$$ 

```{r}
# Calculate squared covariance between r and rows of G
covs <- rep(0, M)
for (i in 1:M) {
  covs[i] <- cov(r, G[i,])^2
}

# Compute numerator 
Num <- mean(covs)

# Calculate variance of rows of G
vars <- rep(0, M)
for (i in 1:M) {
  vars[i] <- var(G[i,])
}

# Compute numerator 
Dom <- mean(vars)

# Compute H 
HCov <- Num / Dom
print(HCov)
```

Direct Estimator of $H$

$$\hat{H} = \frac{1}{M(L-1)^2} r^\top \mathbf{G}^\top \mathbf{G} r$$
```{r}
# Compute estimator directly 
hatH <- (1/(M * (L-1)^2)) * (t(r) %*% t(G) %*% G %*% r)
hatH
```

$F_{Gr}$ Estimator of $H$

$$F_{Gr} = \frac{1}{\sqrt{(L - 1)}} Gr$$
$$\hat{H} = \frac{1}{M (1 -L)} F_{Gr}^\top F_{Gr}$$
```{r}
# Compute FGr 
FGr <- (1/sqrt(L-1)) * (G %*% r)

# Check variance of FGr
var(FGr)

# Compute estimator via FGr
HFGr <- (1/(M * (L-1))) * (t(FGr) %*% FGr)
HFGr
```

Under the null hypothesis of no overlap $E[\hat{H}] \approx 1/L$

```{r}
run_sim_random <- function(L, M) {
  
  # Random r 
  r <- runif(L)
  
  # Make Genotypes
  G <- matrix(NA, nrow = M, ncol = L)
  for (l in 1:L) {
    G[,l] <- runif(M, 0, 1)
  }
  
  # Scale vectors
  r <- scale(r)
  G <- scale(G)
  
  # Compute estimator directly 
  hatH <- (1/(M * (L-1)^2)) * t(r) %*% t(G) %*% G %*% r
  
  return(hatH)
} 
```

L = 200 
```{r}
nums <- rep(0, 500)
for(i in 1:500) {
  nums[i] <- run_sim_random(200, 1000)
}

print(paste0("Expection is ", mean(nums)))
print(paste0("Expection should be ", 1/200))
```

L = 1000 
```{r}
nums <- rep(0, 500)
for(i in 1:500) {
  nums[i] <- run_sim_random(1000, 1000)
}

print(paste0("Expection is ", mean(nums)))
print(paste0("Expection should be ", 1/1000))
```

L = 2000 
```{r}
nums <- rep(0, 500)
for(i in 1:500) {
  nums[i] <- run_sim_random(2000, 1000)
}

print(paste0("Expection is ", mean(nums)))
print(paste0("Expection should be ", 1/2000))
```

## Non-overlapping 4 pop model 

### In R

```{r}
L <- 2000

# Define the file paths
pgen_file <- "../../Downloads/genos.pgen"
pvar_file <- "../../Downloads/genos.pvar"
psam_file <- "../../Downloads/genos.psam"

# Load the PGEN file
pvar <- NewPvar(pvar_file)
pgen <- NewPgen(pgen_file)

# Get L IDs
IDs <- sapply(1:L, function(i) GetVariantId(pvar, i))

# Get GWAS and test panel IDs
testIDs <- fread("~/Downloads/ids.test")
gwasIDs <- fread("~/Downloads/ids.gwas")

# Read in psam file and get index
psam_data <- fread(psam_file)
sample_ids <- psam_data$IID

test_indices <- match(testIDs$IID, sample_ids)
test_indices <- test_indices[!is.na(test_indices)] 

gwas_indices <- match(gwasIDs$IID, sample_ids)
gwas_indices <- gwas_indices[!is.na(gwas_indices)] 

# Read in genotype matrices
geno_mat <- ReadList(pgen, variant_subset = 1:L, meanimpute = FALSE)
X_mat <- geno_mat[test_indices, , drop = FALSE]
G_mat <- geno_mat[gwas_indices, , drop = FALSE]

# Get r 
pC <- colMeans(X_mat[1:500,])
pD <- colMeans(X_mat[501:1000,])
r <- pC - pD

# Scale 
r <- scale(r)
G <- scale(G_mat)
G[is.na(G)] <- 0
```

Direct Estimator of $H$

$$\hat{H} = \frac{1}{M(L-1)^2} r^\top \mathbf{G}^\top \mathbf{G} r$$
```{r}
# Compute estimator directly 
hatH <- (1/(M * (L-1)^2)) * (t(r) %*% t(G) %*% G %*% r)
hatH
```

$F_{Gr}$ Estimator of $H$

$$F_{Gr} = \frac{1}{\sqrt{(L - 1)}} Gr$$
$$\hat{H} = \frac{1}{M (1 -L)} F_{Gr}^\top F_{Gr}$$
```{r}
# Compute FGr
FGr_raw <- (G %*% r)
FGr <- (1/sqrt(L-1)) * (G %*% r)

# Compute variance of FGr
var(FGr)

# Compute estimator
HFGr <- (1/(M * (L-1))) * (t(FGr) %*% FGr)
HFGr
1/L
```

### Via plink 

Compute H hat via plink 
```{r}
# Format R file 
dfr <- cbind(IDs,rep("T", length(r)), r)
dfr <- as.data.frame(dfr) 
fwrite(dfr, "~/Downloads/dfr.txt", quote = F, row.names = F, sep = "\t")

# Run plink command 
plink_cmd <- paste0("~/Downloads/plink2 --pfile ~/Downloads/genos --keep ~/Downloads/ids.gwas --score ~/Downloads/dfr.txt variance-standardize header-read cols=dosagesum,scoresums  --out ~/Downloads/FGr_plink")
system(plink_cmd)

# Read in results
plinkFGr_raw <- fread("~/Downloads/FGr_plink.sscore")
plinkFGr <- (1/sqrt(L-1)) * plinkFGr_raw$V3_SUM

# Compare FGr estimates
plot(sort(plinkFGr),sort(FGr))

# Compute var(FGr)
var(plinkFGr)

# Compute H 
HPlink<- (1/(M * (L-1))) * (t(plinkFGr) %*% plinkFGr)
HPlink
```

## Subset of UKBB genotypes and random r's

### Via R

```{r}
L <- 2000

# Define the file paths
pgen_file <- "../../Downloads/ukb_imp_chr21_v3.pgen"
pvar_file <- "../../Downloads/ukb_imp_chr21_v3.pvar"
psam_file <- "../../Downloads/ukb_imp_chr21_v3.psam"

# Load the PGEN file
pvar <- NewPvar(pvar_file)
pgen <- NewPgen(pgen_file)

# Get L IDs
IDs <- sapply(1:L, function(i) GetVariantId(pvar, i))

# Read in genotypes 
geno_mat <- ReadList(pgen, variant_subset = 1:L, meanimpute = FALSE)

# Simulate r 
r <- runif(L)

# Scale 
r <- scale(r)
G <- scale(geno_mat)
G[is.na(G)] <- 0
```

```{r}
# Compute estimator directly 
hatH <- (1/(M * (L-1)^2)) * (t(r) %*% t(G) %*% G %*% r)
hatH
```

```{r}
# Compute FGr
FGr_raw <- (G %*% r)
FGr <- (1/sqrt(L-1)) * (G %*% r)

# Compute variance of FGr
var(FGr)

# Compute estimator
HFGr <- (1/(M * (L-1))) * (t(FGr) %*% FGr)
HFGr
1/L
```
### Via Plink

Compute H hat via plink 
```{r}
# Format R file 
dfPVar <- fread(pvar_file) %>% select("ID", "ALT")
dfr <- cbind(dfPVar, r)
dfr <- as.data.frame(dfr) 
fwrite(dfr, "~/Downloads/dfr.txt", quote = F, row.names = F, sep = "\t")

# Run plink command 
plink_cmd <- paste0("~/Downloads/plink2 --pfile ~/Downloads/ukb_imp_chr21_v3 --score ~/Downloads/dfr.txt variance-standardize header-read cols=dosagesum,scoresums  --out ~/Downloads/FGr_plink")
system(plink_cmd)

# Read in results
plinkFGr_raw <- fread("~/Downloads/FGr_plink.sscore")
plinkFGr <- (1/sqrt(L-1)) * plinkFGr_raw$V1_SUM

# Compare FGr estimates
plot(sort(plinkFGr),sort(FGr))

# Compute var(FGr)
var(plinkFGr)

# Compute H 
HPlink<- (1/(M * (L-1))) * (t(plinkFGr) %*% plinkFGr)
HPlink
```

## Subset of UKBB genotypes and random r's done in two blocks 

```{r}
# Format R file 
dfPVar <- fread(pvar_file) %>% select("ID", "ALT")
dfr <- cbind(dfPVar, r)
dfr <- as.data.frame(dfr) 
fwrite(dfr[1:1000,], "~/Downloads/dfr1.txt", quote = F, row.names = F, sep = "\t")
fwrite(dfr[1001:2000,], "~/Downloads/dfr2.txt", quote = F, row.names = F, sep = "\t")

# Run plink command 
plink_cmd <- paste0("~/Downloads/plink2 --pfile ~/Downloads/ukb_imp_chr21_v3 --score ~/Downloads/dfr1.txt variance-standardize header-read cols=dosagesum,scoresums  --out ~/Downloads/FGr_plink1")
system(plink_cmd)

# Read in results
plinkFGr1_raw <- fread("~/Downloads/FGr_plink1.sscore")
plinkFGr1 <- (1/sqrt(999)) * plinkFGr1_raw$V1_SUM
var(plinkFGr1)

# Run plink command 
plink_cmd <- paste0("~/Downloads/plink2 --pfile ~/Downloads/ukb_imp_chr21_v3 --score ~/Downloads/dfr2.txt variance-standardize header-read cols=dosagesum,scoresums  --out ~/Downloads/FGr_plink2")
system(plink_cmd)

# Read in results
plinkFGr2_raw <- fread("~/Downloads/FGr_plink2.sscore")
plinkFGr2 <- (1/sqrt(999)) * plinkFGr2_raw$V1_SUM
var(plinkFGr2)

# Combine 
plinkFGrT2 <- (plinkFGr1_raw$V1_SUM + plinkFGr2_raw$V1_SUM) * (1/sqrt(L-1))
var(plinkFGrT2)
plot(sort(plinkFGrT2),sort(FGr))

# Compare to plink no blocks
plot(sort(plinkFGrT2),sort(plinkFGr))

# Compute H 
HPlink<- (1/(M * (L-1))) * (t(plinkFGrT2) %*% plinkFGrT2)
HPlink
```

```{r}
dfSmall <- fread("~/Downloads/small.txt")
colnames(dfSmall) <- c("varRaw","varScale",	"H",	"1/L",	"1/M",	"1/N",	"L")

ggplot(dfSmall, aes(x = `1/L`, y = H)) + geom_point() + scale_x_log10() + scale_y_log10() + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+geom_hline(yintercept = 0.001, color = "purple")+geom_hline(yintercept = 0.000547, color = "green")

ggplot(dfSmall, aes(x = L, y = varScale)) + geom_point() + scale_x_log10() + geom_hline(yintercept = 1, color = "red")
```

```{r}
dfMedium <- fread("~/Downloads/medium.txt")
colnames(dfMedium) <- c("varRaw","varScale",	"H",	"1/L",	"1/M",	"1/N",	"L")

ggplot(dfMedium, aes(x = `1/L`, y = H)) + geom_point() + scale_x_log10() + scale_y_log10() + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+geom_hline(yintercept = 1e-04, color = "purple")+geom_hline(yintercept = 0.000547, color = "green")

ggplot(dfMedium, aes(x = L, y = varScale)) + geom_point() + scale_x_log10() + geom_hline(yintercept = 1, color = "red")
```

```{r}
dfLarge <- fread("~/Downloads/large.txt")
colnames(dfLarge) <- c("varRaw","varScale",	"H",	"1/L",	"1/M",	"1/N",	"L")

ggplot(dfLarge, aes(x = `1/L`, y = H)) + geom_point() + scale_x_log10() + scale_y_log10() + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+geom_hline(yintercept = 1e-05, color = "purple")+geom_hline(yintercept = 0.000547, color = "green")

ggplot(dfLarge, aes(x = L, y = varScale)) + geom_point() + scale_x_log10() + geom_hline(yintercept = 1, color = "red")
```

```{r}
dfSmall$M <- 1/dfSmall$`1/M`
dfMedium$M <- 1/dfMedium$`1/M`
dfLarge$M <- 1/dfLarge$`1/M`
dfAll <- rbind(dfSmall, dfMedium, dfLarge)

dfAll$M <- as.character(dfAll$M)
ggplot(dfAll, aes(x = L, y = varScale, color = M)) + geom_point() + scale_x_log10() + geom_hline(yintercept = 1, color = "red") + ylab("Variance of FGr (should be 1)")
```
```{r}
ggplot(dfAll, aes(x = `1/L`, y = H, color = M)) + geom_point() + scale_x_log10() + scale_y_log10() + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +geom_hline(yintercept = 1e-4, color = "#00BA38") + geom_hline(yintercept = 1e-3, color = "#F8766D") + geom_hline(yintercept = 1e-5, color = "#619CFF")
```



```{r}
df <- fread("~/Downloads/block_L-pruneall.txt")
df$r <- runif(nrow(df))
df$ID <- runif(nrow(df)) *100

dfSum <- df %>% group_by(block) %>% summarise(nSNP = n())

rep = 1
dfNew <- df %>%mutate(r = c(tail(r, rep * 1000), head(r, -rep *1000)))
```


