---
title: "FGrPlot"
output: html_document
date: "2025-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(latex2exp)
library("gridExtra")
library(grid)
library(scales)
```


## Plot all Error 

```{r}
# Overlap data
tmp1 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt")
dfAllOverlap <- rbind(tmp1, tmp2)

# Error data
tmp1 <- fread("../plots/error/error_HGDP1KG.txt")
tmp2 <- fread("../plots/error/error_InUKBB.txt")
dfError <- rbind(tmp1, tmp2)

# Combine
df <- inner_join(dfError, dfAllOverlap)

# Format Data
tmp1 <- df %>% filter(gwas %in% c("e5", "e10", "e50", "e200")) %>% separate(gwas, into = c("tmp", "gwas"), sep = "e") %>% select(-tmp)
tmp1$gwas <- factor(tmp1$gwas, levels = c( "5", "10", "50", "200")) 
tmp2 <- df %>% filter(gwas %in% c("WBS-R", "ALL-R")) %>% separate(gwas, into = c("gwas", "tmp"), sep = "-") %>% select(-tmp)
tmp3 <-  df %>% filter(gwas %in% c("WBS", "ALL"))
df <- rbind(tmp1, tmp2, tmp3)
df$gwas <- factor(df$gwas, levels = c("WBS", "5", "10", "50", "200", "ALL"))
```

```{r}
pl <- df %>% ggplot(aes(x = H, y = error, color = gwas)) + geom_point(size = 2) + scale_x_log10() + scale_y_log10() + xlab(TeX("Proportion of variance explained ($\\hat{H}$)")) + ylab(TeX("Error ($\\hat{e}_{f}$)")) + theme_bw(base_size = 12) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), labels = c(expression("WBS"),expression(5 * epsilon), expression(10 * epsilon),expression(50 * epsilon),expression(200 * epsilon), expression("ALL")))  + theme(panel.grid.minor = element_line(linetype = "dashed")) +  guides(color = guide_legend(title = "GWAS", label = scales::label_parse()))
#ggsave("~/Desktop/Error.png", pl, height = 4, width = 6)
```

## Plot R2 + Ratio Even/Odd

```{r}
tmp1 <- fread("../plots/error/R2EO_HGDP1KG.txt")
tmp2 <- fread("../plots/error/R2EO_InUKBB.txt")
df <- rbind(tmp1, tmp2)

# Format Data
tmp1 <- df %>% filter(gwas %in% c("e5", "e10", "e50", "e200")) %>% separate(gwas, into = c("tmp", "gwas"), sep = "e") %>% select(-tmp)
tmp1$gwas <- factor(tmp1$gwas, levels = c( "5", "10", "50", "200")) 
tmp2 <- df %>% filter(gwas %in% c("WBS-R", "ALL-R")) %>% separate(gwas, into = c("gwas", "tmp"), sep = "-") %>% select(-tmp)
tmp3 <-  df %>% filter(gwas %in% c("WBS", "ALL"))
df <- rbind(tmp1, tmp2, tmp3)
df$gwas <- factor(df$gwas, levels = c("WBS", "5", "10", "50", "200", "ALL"))

# Shorten Names
rep_str = c("England" = "ENG","Wales" = "WAL", "Scotland" = "SCT","NorthernIreland" = "NI","RepublicOfIrelend" = "RoI" )
df$contrasts <- str_replace_all(df$contrasts, rep_str)
```

Plot Signal 
```{r}
pl <- df %>% ggplot(aes(x = Signal,y = Ratio, color = gwas)) + geom_point() + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), labels = c(expression("WBS"),expression(5 * epsilon), expression(10 * epsilon),expression(50 * epsilon),expression(200 * epsilon), expression("ALL"))) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + geom_vline(xintercept = 0, color = "red", linetype = "dashed") +  geom_vline(xintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 12) + theme(panel.grid.minor = element_line(linetype = "dashed")) +  guides(color = guide_legend(title = "GWAS", label = scales::label_parse())) +  xlab(TeX("Signal ($\\hat{\\gamma}_{f}$)")) + ylab(TeX("Signal captured by PCs ($V_{k}$)"))

ggsave("~/Desktop/SignalRatio.png", pl, height = 6, width = 10)
```

Plot R2 for each set of contrasts

HGDP - continental 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))  

plCont <- dfPlot %>% filter(PC <= 20) %>% ggplot(aes(x = PC, y = R2, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(aes(yintercept = Signal, color = gwas), linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Variance captured by PCs ($R^2_{k}$)"))
```

HGDP - lat/long 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eurasia-lat", "eurasia-long", "eur-lat", "eur-long", "sdi-eur"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eurasia-lat", "eurasia-long","eur-lat", "eur-long", "sdi-eur"))

plLat <- dfPlot %>% filter(PC <= 20) %>% ggplot(aes(x = PC, y = R2, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(aes(yintercept = Signal, color = gwas), linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Variance captured by PCs ($R^2_{k}$)"))
```

InUKBB 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels =c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))  

plInUKBB <- dfPlot %>% filter(PC <= 20) %>% ggplot(aes(x = PC, y = R2, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(aes(yintercept = Signal, color = gwas), linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), name= TeX("GWAS"),  labels = c(expression("WBS"),expression(5 * epsilon), expression(10 * epsilon),expression(50 * epsilon),expression(200 * epsilon), expression("ALL"))) + theme( legend.position = "bottom", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Variance captured by PCs ($R^2_{k}$)")) + ylim(0,1) +
  guides(color = guide_legend(nrow = 1))
```

Combine R2 Figures 
```{r}
# Define label grobs
label_a <- textGrob("a", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_b <- textGrob("b", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_c <- textGrob("c", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))


# Wrap other plots with labels
row_a <- arrangeGrob(plCont, top = label_a)
row_b <- arrangeGrob(plLat, top = label_b)
row_c <- arrangeGrob(plInUKBB, top = label_c)

# Final arrangement
pl <- grid.arrange( row_a, row_b, row_c, ncol = 1, heights = c(1, 1, 1.3))
ggsave("~/Desktop/SupR2.png", pl, height = 10, width = 12)
```


Plot Ratio for each set of contrasts

HGDP - continental 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))  

plCont <- dfPlot %>% filter(PC <= 20) %>% ggplot(aes(x = PC, y = Ratio, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Fraction of signal captured PCs ($V_{k}$)")) + geom_errorbar(aes(x = PC, ymin = lc, ymax = uc), width = 0.6)
plCont
```

HGDP - lat/long 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eurasia-lat", "eurasia-long", "eur-lat", "eur-long", "sdi-eur"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eurasia-lat", "eurasia-long","eur-lat", "eur-long", "sdi-eur"))

plLat <- dfPlot %>% filter(PC <= 20) %>% ggplot(aes(x = PC, y = Ratio, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Fraction of signal captured PCs ($V_{k}$)")) + geom_errorbar(aes(x = PC, ymin = lc, ymax = uc), width = 0.6)
```


InUKBB 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels =c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))  

plInUKBB <- dfPlot %>% filter(PC <= 20) %>% ggplot(aes(x = PC, y = Ratio, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("orchid1","purple1", "darkorange", "skyblue2", "seagreen", "seagreen1"), name= TeX("GWAS")) + theme( legend.position = "bottom", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Fraction of signal captured PCs ($V_{k}$)")) +
  guides(color = guide_legend(nrow = 1)) + ylim(0,1) + geom_errorbar(aes(x = PC, ymin = lc, ymax = uc), width = 0.6)
```

Combine Ratio Figures 
```{r}
# Define label grobs
label_a <- textGrob("a", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_b <- textGrob("b", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_c <- textGrob("c", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))


# Wrap other plots with labels
row_a <- arrangeGrob(plCont, top = label_a)
row_b <- arrangeGrob(plLat, top = label_b)
row_c <- arrangeGrob(plInUKBB, top = label_c)

# Final arrangement
pl <- grid.arrange( row_a, row_b, row_c, ncol = 1, heights = c(1, 1, 1.3))
ggsave("~/Desktop/Ratio.png", pl, height = 10, width = 12)
```

## Plot R2 + Ratio Even/Odd Rare PCA

```{r}
tmp1 <- fread("../plots/error/rare_R2EO_HGDP1KG.txt")
tmp2 <- fread("../plots/error/rare_R2EO_InUKBB.txt")
df <- rbind(tmp1, tmp2)

# Format Data
tmp1 <- df %>% filter(gwas %in% c("e5", "e10", "e50", "e200")) %>% separate(gwas, into = c("tmp", "gwas"), sep = "e") %>% select(-tmp)
tmp1$gwas <- factor(tmp1$gwas, levels = c( "5", "10", "50", "200")) 
tmp2 <- df %>% filter(gwas %in% c("WBS-R", "ALL-R")) %>% separate(gwas, into = c("gwas", "tmp"), sep = "-") %>% select(-tmp)
tmp3 <-  df %>% filter(gwas %in% c("WBS", "ALL"))
df <- rbind(tmp1, tmp2, tmp3)
df$gwas <- factor(df$gwas, levels = c("WBS", "5", "10", "50", "200", "ALL"))

# Shorten Names
rep_str = c("England" = "ENG","Wales" = "WAL", "Scotland" = "SCT","NorthernIreland" = "NI","RepublicOfIrelend" = "RoI" )
df$contrasts <- str_replace_all(df$contrasts, rep_str)
```

Plot Signal 
```{r}
pl <- df  %>%  ggplot(aes(x = Signal,y = Ratio, color = gwas)) + geom_point() + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), labels = c(expression(5 * epsilon), expression(10 * epsilon),expression(50 * epsilon),expression(200 * epsilon))) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + geom_vline(xintercept = 0, color = "red", linetype = "dashed") +  geom_vline(xintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 12) + theme(panel.grid.minor = element_line(linetype = "dashed")) +  guides(color = guide_legend(title = "GWAS", label = scales::label_parse())) +  xlab(TeX("Signal ($\\hat{\\gamma}_{f}$)")) + ylab(TeX("Signal captured by PCs ($V_{k}$)"))
pl
ggsave("~/Desktop/SignalRatio_Rare.png", pl, height = 6, width = 10)
```

Plot R2 for each set of contrasts

HGDP - continental 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))  

plCont <- dfPlot %>% filter(PC <= 40) %>% ggplot(aes(x = PC, y = R2, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(aes(yintercept = Signal, color = gwas), linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Variance captured by PCs ($R^2_{k}$)"))
```

HGDP - lat/long 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eurasia-lat", "eurasia-long", "eur-lat", "eur-long", "sdi-eur"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eurasia-lat", "eurasia-long","eur-lat", "eur-long", "sdi-eur"))

plLat <- dfPlot %>% filter(PC <= 40) %>% ggplot(aes(x = PC, y = R2, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(aes(yintercept = Signal, color = gwas), linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Variance captured by PCs ($R^2_{k}$)"))
```

InUKBB 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels =c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))  

plInUKBB <- dfPlot %>% filter(PC <= 40) %>% ggplot(aes(x = PC, y = R2, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(aes(yintercept = Signal, color = gwas), linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), name= TeX("GWAS"),  labels = c(expression(5 * epsilon), expression(10 * epsilon),expression(50 * epsilon),expression(200 * epsilon)))+ theme( legend.position = "bottom", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Variance captured by PCs ($R^2_{k}$)")) + ylim(0,1) +
  guides(color = guide_legend(nrow = 1))
```

Combine R2 Figures 
```{r}
# Define label grobs
label_a <- textGrob("a", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_b <- textGrob("b", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_c <- textGrob("c", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))


# Wrap other plots with labels
row_a <- arrangeGrob(plCont, top = label_a)
row_b <- arrangeGrob(plLat, top = label_b)
row_c <- arrangeGrob(plInUKBB, top = label_c)

# Final arrangement
pl <- grid.arrange( row_a, row_b, row_c, ncol = 1, heights = c(1, 1, 1.3))
ggsave("~/Desktop/SupR2_rare.png", pl, height = 10, width = 12)
```


Plot Ratio for each set of contrasts

HGDP - continental 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr"))  

plCont <- dfPlot %>% filter(PC <= 40) %>% ggplot(aes(x = PC, y = Ratio, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Fraction of signal captured PCs ($V_{k}$)")) + geom_errorbar(aes(x = PC, ymin = lc, ymax = uc), width = 0.6)
plCont
```

HGDP - lat/long 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("eurasia-lat", "eurasia-long", "eur-lat", "eur-long", "sdi-eur"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels = c("eurasia-lat", "eurasia-long","eur-lat", "eur-long", "sdi-eur"))

plLat <- dfPlot %>% filter(PC <= 40) %>% ggplot(aes(x = PC, y = Ratio, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), name= TeX("Sampling distance ($\\epsilon$)")) + theme( legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Fraction of signal captured PCs ($V_{k}$)")) + geom_errorbar(aes(x = PC, ymin = lc, ymax = uc), width = 0.6)
```

InUKBB 
```{r}
dfPlot <- df %>% filter(contrasts %in%c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))
dfPlot$contrasts <- factor(dfPlot$contrasts, levels =c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL"))  

plInUKBB <- dfPlot %>% filter(PC <= 40) %>% ggplot(aes(x = PC, y = Ratio, color = gwas)) + geom_point() + geom_line() + facet_wrap(~contrasts, nrow = 1) + geom_hline(yintercept = 1, color = "red", linetype = "dashed") + theme_bw(base_size = 10) + scale_color_manual(values = c("purple1", "darkorange", "skyblue2", "seagreen"), name= TeX("GWAS")) + theme( legend.position = "bottom", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab("PC") + ylab(TeX("Fraction of signal captured PCs ($V_{k}$)")) +
  guides(color = guide_legend(nrow = 1)) + ylim(0,1) + geom_errorbar(aes(x = PC, ymin = lc, ymax = uc), width = 0.6)
```

Combine Ratio Figures 
```{r}
# Define label grobs
label_a <- textGrob("a", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_b <- textGrob("b", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_c <- textGrob("c", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))


# Wrap other plots with labels
row_a <- arrangeGrob(plCont, top = label_a)
row_b <- arrangeGrob(plLat, top = label_b)
row_c <- arrangeGrob(plInUKBB, top = label_c)

# Final arrangement
pl <- grid.arrange( row_a, row_b, row_c, ncol = 1, heights = c(1, 1, 1.3))
ggsave("~/Desktop/Ratio_rare.png", pl, height = 10, width = 12)
```



## Plot R2 by Chr 

```{r}
tmp1 <- fread("../plots/error/R2Chr_HGDP1KG.txt")
tmp2 <- fread("../plots/error/R2Chr_InUKBB.txt")
df <- rbind(tmp1, tmp2)

# Format Data
tmp1 <- df %>% filter(gwas %in% c("e5", "e10", "e50", "e200")) %>% separate(gwas, into = c("tmp", "gwas"), sep = "e") %>% select(-tmp)
tmp1$gwas <- factor(tmp1$gwas, levels = c( "5", "10", "50", "200")) 
tmp2 <- df %>% filter(gwas %in% c("WBS-R", "ALL-R")) %>% separate(gwas, into = c("gwas", "tmp"), sep = "-") %>% select(-tmp)
tmp3 <-  df %>% filter(gwas %in% c("WBS", "ALL"))
df <- rbind(tmp1, tmp2, tmp3)
df$gwas <- factor(df$gwas, levels = c("WBS", "5", "10", "50", "200", "ALL"))

# Shorten Names
rep_str = c("England" = "ENG","Wales" = "WAL", "Scotland" = "SCT","NorthernIreland" = "NI","RepublicOfIrelend" = "RoI" )
df$contrasts <- str_replace_all(df$contrasts, rep_str)
```

#### Signal vs Number of Chromosomes used to calculate it 

```{r}
df %>% filter(PC == "1" ) %>% filter(contrasts %in%c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr")) %>% mutate("FGr_chrs" = 22 - pca_chrs) %>% ggplot(aes(x = pca_chrs, y = Signal, color = gwas)) + geom_point() + facet_wrap(~contrasts)
```


```{r}
df %>% filter(PC == "1" ) %>% filter(contrasts %in%c("eurasia-lat", "eurasia-long", "eur-lat", "eur-long", "sdi-eur")) %>% mutate("FGr_chrs" = 22 - pca_chrs) %>% ggplot(aes(x = pca_chrs, y = Signal, color = gwas)) + geom_point() + facet_wrap(~contrasts)
```


```{r}
df %>% filter(PC == "1" ) %>% filter(contrasts %in%c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL")) %>% mutate("FGr_chrs" = 22 - pca_chrs) %>% ggplot(aes(x = pca_chrs, y = Signal, color = gwas)) + geom_point() + facet_wrap(~contrasts)
```

### Ratio vs Chr split


```{r}
df %>% filter(PC == "40") %>% filter(contrasts %in%c("eas-nfe","eas-sas","eas-afr", "eas-amr", "nfe-sas","nfe-afr", "nfe-amr","sas-afr", "sas-amr", "afr-amr")) %>% mutate("FGr_chrs" = 22 - pca_chrs) %>% ggplot(aes(x = pca_chrs, y = Ratio, color = gwas)) + geom_point() + facet_wrap(~contrasts) + ylim(0,1)
```

```{r}
df %>% filter(PC == "40") %>% filter(contrasts %in%c("eurasia-lat", "eurasia-long", "eur-lat", "eur-long", "sdi-eur")) %>% mutate("FGr_chrs" = 22 - pca_chrs) %>% ggplot(aes(x = pca_chrs, y = Ratio, color = gwas)) + geom_point() + facet_wrap(~contrasts) 
```


```{r}
df %>% filter(PC == "40") %>% filter(contrasts %in%c("ENG-NI", "ENG-RoI", "ENG-SCT", "ENG-WAL", "NI-RoI", "NI-SCT", "NI-WAL", "RoI-SCT", "RoI-WAL", "SCT-WAL")) %>% mutate("FGr_chrs" = 22 - pca_chrs) %>% ggplot(aes(x = pca_chrs, y = Ratio, color = gwas)) + geom_point() + facet_wrap(~contrasts)
```


