---
title: "QPlot"
output: html_document
date: "2025-06-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(latex2exp)
library("gridExtra")
library(grid)
library(scales)
library(ggtext)
```



## Resampled SNPs

```{r}
# Read in all Qs 
tmp1 <- fread("../plots/pga_test_resampled/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/pga_test_resampled/HGDP1KG/WBS/q_results.txt")
tmp3 <- fread("../plots/pga_test_resampled/InUKBB/WBS-R/q_results.txt") %>% mutate(gwas = "WBS")
tmp4 <- fread("../plots/pga_test_resampled/InUKBB/ALL-R/q_results.txt") %>% mutate(gwas = "ALL")
dfQ <- rbind(tmp1, tmp2, tmp3, tmp4)

# Read in H values
tmp1 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt") %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfH <- rbind(tmp1, tmp2) %>% select(H, L, gwas, contrasts, dataset) 

# Read in ratio values 
tmp1 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(dataset = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset)
tmp2 <- fread("../plots/error/common_rare_R2EO_InUKBB.txt") %>% mutate(dataset = "InUKBB") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset) %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfRatio <- rbind(tmp1, tmp2)

# Combine
df <- inner_join(dfQ, dfH)
df <- inner_join(df, dfRatio)
```


```{r}
# Multiple Testing correction w/in each gwas/covar/gytpe/threshold
df <- df %>% group_by(gwas, covar, gtype, threshold) %>% mutate(adjuP = p.adjust(pval, method = "bonferroni")) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F)) %>% mutate(HPrime = H * (1 - Ratio))
```

## Q^2 vs H

```{r}
df %>% filter() %>% ggplot(aes(x = H, y = q^2, color = covar)) + 
  geom_jitter(data = ~ subset(., covar == "nopcs"), 
              aes(color = covar), alpha = 0.5) +
  geom_jitter(data = ~ subset(., covar == "pcs"), 
              aes(color = covar), alpha = 0.5) +
  scale_x_log10() + 
  scale_y_log10() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) + 
  scale_color_manual(
  name = "PCs included",  # Legend title
  values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
  labels = c("nopcs" = "none", "pcs" = "40 common + 40 rare"))+
  theme_bw(base_size = 12) + 
  theme(legend.position = "bottom", strip.background=element_rect(colour="black",fill="white")) + 
  xlab(TeX("$\\hat{H}")) + 
  ylab(TeX("$\\hat{q}^2"))
```

```{r}
library(scales)  # for alpha()

# Create q^2 and reshape to long format
df_long <- df %>%
  mutate(q2 = q^2) %>%
  pivot_longer(cols = c(H, HPrime), names_to = "x_type", values_to = "xval")

ggplot(df_long, aes(x = xval, y = q2, color = covar)) + 
  
  # Plot nopcs points first (go underneath)
  geom_jitter(data = subset(df_long, x_type == "H" & covar == "nopcs"), alpha = 0.05) +
  
  # Plot pcs points second (go on top)
  geom_jitter(data = subset(df_long, x_type == "H" & covar == "pcs"), alpha = 0.05) +
  geom_jitter(data = subset(df_long, x_type == "HPrime" & covar == "pcs"), alpha = 0.5) +
  
  # Faint smooths for H (use alpha-colored lines)
  geom_smooth(
    data = subset(df_long, x_type == "H" & covar == "pcs"),
    aes(x = xval, y = q2), method = "lm", se = FALSE,
    color = alpha("darkred", 0.2)
  ) +
  geom_smooth(
    data = subset(df_long, x_type == "H" & covar == "nopcs"),
    aes(x = xval, y = q2), method = "lm", se = FALSE,
    color = alpha("goldenrod2", 0.2)
  ) +
  
  # Colored smooths for HPrime (appear in legend)
  geom_smooth(
    data = subset(df_long, x_type == "HPrime"),
    aes(x = xval, y = q2, color = covar), method = "lm", se = FALSE
  ) +
  
  scale_x_log10() + 
  scale_y_log10() + 
  
  facet_grid(
    vars(gtype), vars(threshold),
    labeller = labeller(
      gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
      threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
    )
  ) +
  
  scale_color_manual(
    name = "PCs included",
    values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
    labels = c("nopcs" = "none", "pcs" = "40 common + 40 rare")
  ) +
  
  theme_bw(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  xlab(TeX("$\\hat{H}$ or $\\hat{H}'$")) + 
  ylab(TeX("$\\hat{q}^2$"))

```


```{r}
df$H_ratio <- df$H / df$HPrime

df %>% filter() %>% ggplot(aes(x = H_ratio, y = q^2, color = covar)) + 
  geom_jitter(data = ~ subset(., covar == "nopcs"), 
              aes(color = covar), alpha = 0.5) +
  geom_jitter(data = ~ subset(., covar == "pcs"), 
              aes(color = covar), alpha = 0.5) +
  scale_x_log10() + 
  scale_y_log10() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) + 
  scale_color_manual(
  name = "PCs included",  # Legend title
  values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
  labels = c("nopcs" = "none", "pcs" = "40 common + 40 rare"))+
  theme_bw(base_size = 12) + 
  theme(legend.position = "bottom", strip.background=element_rect(colour="black",fill="white")) + 
  xlab(TeX("$\\frac{H}{H'}")) + 
  ylab(TeX("$\\hat{q}^2"))
```




## Finding Significant Results 



```{r}
dfPCs <- df %>% filter(covar == "pcs")
dfNoPCs <- df %>% filter(covar == "nopcs")

dfPlot <- dfPCs
dfPlot$q_nopcs <- dfNoPCs$q
dfPlot$q_ratio <- (dfPlot$q_nopcs / dfPlot$q)^2
dfPlot$H_ratio <- dfPlot$H / dfPlot$HPrime

minL <- min(dfPlot$L)
maxL <- max(dfPlot$L)

dfPlot %>% ggplot(aes(x = HPrime, y = -log10(adjuP), color = sigP)) +
  geom_jitter() +
  scale_x_log10() + theme_bw() + 
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) +
  geom_vline(xintercept = 1/maxL, color = "red", linetype = "dashed") + geom_vline(xintercept = 1/minL,color = "red", linetype = "dashed")
```



## Not-Resampled SNPs

```{r}
# Read in all Qs 
tmp1 <- fread("../plots/pga_test/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/pga_test/HGDP1KG/WBS/q_results.txt")
tmp3 <- fread("../plots/pga_test/InUKBB/WBS-R/q_results.txt") %>% mutate(gwas = "WBS")
tmp4 <- fread("../plots/pga_test/InUKBB/ALL-R/q_results.txt") %>% mutate(gwas = "ALL")
dfQ <- rbind(tmp1, tmp2, tmp3, tmp4)

# Read in H values
tmp1 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt") %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfH <- rbind(tmp1, tmp2) %>% select(H, L, gwas, contrasts, dataset) 

# Read in ratio values 
tmp1 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(dataset = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset)
tmp2 <- fread("../plots/error/common_rare_R2EO_InUKBB.txt") %>% mutate(dataset = "InUKBB") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset) %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfRatio <- rbind(tmp1, tmp2)

# Combine
df <- inner_join(dfQ, dfH)
df <- inner_join(df, dfRatio)
```


```{r}
# Multiple Testing correction w/in each gwas/covar/gytpe/threshold
df <- df %>% group_by(gwas, covar, gtype, threshold) %>% mutate(adjuP = p.adjust(pval, method = "bonferroni")) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F)) %>% mutate(HPrime = H * (1 - Ratio))

#df %>% summarise(count = n())
```

```{r}
df %>% filter() %>% ggplot(aes(x = H, y = q^2, color = covar)) + 
  geom_jitter(data = ~ subset(., covar == "nopcs"), 
              aes(color = covar), alpha = 0.5) +
  geom_jitter(data = ~ subset(., covar == "pcs"), 
              aes(color = covar), alpha = 0.5) +
  scale_x_log10() + 
  scale_y_log10() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) + 
  scale_color_manual(
  name = "PCs included",  # Legend title
  values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
  labels = c("nopcs" = "none", "pcs" = "40 common + 40 rare"))+
  theme_bw(base_size = 12) + 
  theme(legend.position = "bottom", strip.background=element_rect(colour="black",fill="white")) + 
  xlab(TeX("$\\hat{H}")) + 
  ylab(TeX("$\\hat{q}^2"))
```










## CoB in WBS 

```{r}
df <- fread("../plots/pga_test_resampled/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(type = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```

### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```


## CoB in ALL


```{r}
df <- fread("../plots/pga_test/InUKBB/ALL-R/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_InUKBB.txt") %>% mutate(type = "CoB") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Strict Threshold")
```


Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```


### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```



Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```


## HGDP in WBS


```{r}
df <- fread("../plots/pga_test/HGDP1KG/WBS/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(type = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```


Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```


### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```



Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```

## HGDP in ALL 

```{r}
df <- fread("../plots/pga_test/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(type = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```

### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```

