---
title: "QPlot"
output: html_document
date: "2025-06-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(latex2exp)
library("gridExtra")
library(grid)
library(scales)
library(ggtext)
```



## Resampled SNPs

```{r}
# Read in all Qs 
tmp1 <- fread("../plots/pga_test_resampled/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/pga_test_resampled/HGDP1KG/WBS/q_results.txt")
tmp3 <- fread("../plots/pga_test_resampled/InUKBB/WBS-R/q_results.txt") %>% mutate(gwas = "WBS")
tmp4 <- fread("../plots/pga_test_resampled/InUKBB/ALL-R/q_results.txt") %>% mutate(gwas = "ALL")
dfQ <- rbind(tmp1, tmp2, tmp3, tmp4)

# Read in H values
tmp1 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt") %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfH <- rbind(tmp1, tmp2) %>% select(H, L, gwas, contrasts, dataset) 

# Read in ratio values 
tmp1 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(dataset = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset)
tmp2 <- fread("../plots/error/common_rare_R2EO_InUKBB.txt") %>% mutate(dataset = "InUKBB") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset) %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfRatio <- rbind(tmp1, tmp2)

# Combine
df <- inner_join(dfQ, dfH)
df <- inner_join(df, dfRatio)
```


```{r}
# Multiple Testing correction w/in each gwas/covar/gytpe/threshold
df <- df %>% group_by(gwas, covar, gtype, threshold) %>% mutate(adjuP = p.adjust(pval, method = "bonferroni")) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F)) %>% mutate(HPrime = H * (1 - Ratio))
```

## Q^2 vs H

```{r}
df$H_ratio <- df$H / df$HPrime
plResample <- df %>% filter() %>% ggplot(aes(x = H, y = q^2, color = covar)) + 
  geom_jitter(data = ~ subset(., covar == "nopcs"), 
              aes(color = covar), alpha = 0.5) +
  geom_jitter(data = ~ subset(., covar == "pcs"), 
              aes(color = covar), alpha = 0.5) +
  scale_x_log10() + 
  scale_y_log10() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) + 
  scale_color_manual(
  name = "PCs included",  # Legend title
  values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
  labels = c("nopcs" = "none", "pcs" = "40 common + 40 rare"))+
  theme_bw(base_size = 12) + 
  theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white"),plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size = 8))  + 
  xlab(TeX("$\\hat{H}")) + 
  ylab(TeX("$\\hat{q}^2")) + 
  ggtitle("effect sizes only")

plResample
```

## Not-Resampled SNPs

```{r}
# Read in all Qs 
tmp1 <- fread("../plots/pga_test/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/pga_test/HGDP1KG/WBS/q_results.txt")
tmp3 <- fread("../plots/pga_test/InUKBB/WBS-R/q_results.txt") %>% mutate(gwas = "WBS")
tmp4 <- fread("../plots/pga_test/InUKBB/ALL-R/q_results.txt") %>% mutate(gwas = "ALL")
dfQ <- rbind(tmp1, tmp2, tmp3, tmp4)

# Read in H values
tmp1 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt") %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfH <- rbind(tmp1, tmp2) %>% select(H, L, gwas, contrasts, dataset) 

# Read in ratio values 
tmp1 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(dataset = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset)
tmp2 <- fread("../plots/error/common_rare_R2EO_InUKBB.txt") %>% mutate(dataset = "InUKBB") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas, dataset) %>% mutate(gwas = case_when(gwas == "ALL-R" ~ "ALL",gwas == "WBS-R" ~ "WBS"))
dfRatio <- rbind(tmp1, tmp2)

# Combine
df <- inner_join(dfQ, dfH)
df <- inner_join(df, dfRatio)
```


```{r}
# Multiple Testing correction w/in each gwas/covar/gytpe/threshold
df <- df %>% group_by(gwas, covar, gtype, threshold) %>% mutate(adjuP = p.adjust(pval, method = "bonferroni")) %>% mutate(sigP = case_when(adjuP < 0.025  ~ TRUE, adjuP >= 0.025 ~ F)) %>% mutate(HPrime = H * (1 - Ratio))

df %>% summarise(count = n())
```

```{r}
df$H_ratio <- df$H / df$HPrime

plRaw <- df %>% filter() %>% ggplot(aes(x = H, y = q^2, color = covar)) + 
  geom_jitter(data = ~ subset(., covar == "nopcs"), 
              aes(color = covar), alpha = 0.5) +
  geom_jitter(data = ~ subset(., covar == "pcs"), 
              aes(color = covar), alpha = 0.5) +
  scale_x_log10() + 
  scale_y_log10() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) + 
  scale_color_manual(
  name = "PCs included",  # Legend title
  values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
  labels = c("nopcs" = "no pcs", "pcs" = "40 common + 40 rare"))+
  theme_bw(base_size = 12) + 
  theme(legend.position = "right", strip.background=element_rect(colour="black",fill="white"),plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size = 8)) + 
  xlab(TeX("$\\hat{H}")) + 
  ylab(TeX("$\\hat{q}^2")) + 
  ggtitle("ascertainment effect")
```


### Final Figure Confounding

```{r}
library(gridExtra)
library(grid)

# Define label grobs
label_a <- textGrob("a", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_b <- textGrob("b", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))

# Wrap plots with labels
row_a <- arrangeGrob(plResample, top = label_a)
row_b <- arrangeGrob(plRaw, top = label_b)


# Final arrangement: top_row + c, d, e stacked below
final_plot <- grid.arrange(row_a, row_b,
                           ncol = 2, widths = c(0.72,1))

# Save
ggsave("~/Desktop/ConfoundingFig.png", final_plot, height = 6, width = 13)

```

### Supplemental SNP Number Figure 

```{r}
dfSNPNum <- df %>% filter() %>% ggplot(aes(x =nsnp, y = q^2, color = covar)) + 
  geom_jitter(data = ~ subset(., covar == "nopcs"), 
              aes(color = covar), alpha = 0.5) +
  geom_jitter(data = ~ subset(., covar == "pcs"), 
              aes(color = covar), alpha = 0.5) +
  facet_grid(vars(gtype), vars(threshold),labeller = labeller(
    gtype = c("LR" = "Linear Regression", "LMM" = "LMM"),
    threshold = c("strict" = "Strict Threshold (5e-8)", "loose" = "Liberal Threshold (5e-3)")
  )) + 
  scale_y_log10() +
  scale_color_manual(
  name = "PCs included",  # Legend title
  values = c("nopcs" = "goldenrod2", "pcs" = "darkred"),
  labels = c("nopcs" = "no pcs", "pcs" = "40 common + 40 rare"))+
  theme_bw(base_size = 12) + 
  theme(legend.position = "right", strip.background=element_rect(colour="black",fill="white"),plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size = 8)) + 
  xlab(TeX("Number of SNPs in PGS")) + 
  ylab(TeX("$\\hat{q}^2")) 

# Save
ggsave("~/Desktop/SNPNum.png", dfSNPNum, height = 6, width = 10)
```


## Finding Significant Results 

```{r}
dfPCs <- df %>% filter(covar == "pcs") %>% mutate(HPrime = H * (1 - Ratio), log10P = -1 * log10(pval)) %>% filter(gtype == "LMM")

dfNoPCs <- df %>% filter(covar == "nopcs") %>% mutate(HPrime = H * (1 - Ratio), log10P = -1 * log10(pval)) %>% filter(gtype == "LMM")

dfPCs$q_nopcs <- dfNoPCs$q

sig <- dfPCs %>% filter(sigP == TRUE)

df$sigP <- factor(df$sigP, levels = c(TRUE, FALSE))
```

```{r}
plManhattan <- dfPCs %>% 
  ggplot(aes(x = HPrime, y = log10P, color = sigP)) + 
  geom_point() +
  geom_hline(yintercept = -log10(0.025 / 425),color = "red", linetype = "dashed") + 
  facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(
    threshold = c("strict" = "Strict Threshold (5e-8)", 
                  "loose" = "Liberal Threshold (5e-3)"))) +
  ylab("-log10(P)") +
  xlab(TeX("$\\hat{H}'")) + 
  theme_bw(base_size = 12) +
  theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 8)) +
  scale_color_manual(values = c("gray70", "violet"))
plManhattan
```

SDI - Height
```{r}
pl1 <- df %>% filter(contrasts == "sdi-eur", gtype == "LMM") %>% filter(phenotype == "Standing_Height") %>% mutate(upper = q + 3.85, lower = q - 3.85) %>%  ggplot(aes(x = covar, y = q, color = covar, shape = sigP)) + geom_point(size = 3) +  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5) + theme_bw(base_size = 12) + scale_color_manual(values = c("goldenrod", "darkred")) + theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white")) + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + ylab(TeX("standing height $\\hat{q}$ (sdi - eur)")) + xlab("") +facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(threshold = c("strict" = "Strict Threshold", "loose" = "Liberal Threshold"))) + scale_shape_manual(values = c(8,16))
```

EurLat - Height
```{r}
pl2 <- df %>% filter(contrasts == "eur-lat", gtype == "LMM") %>% filter(phenotype == "Standing_Height") %>% mutate(upper = q + 3.85, lower = q - 3.85) %>%  ggplot(aes(x = covar, y = q, color = covar, shape = sigP)) + geom_point(size = 3) +  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5) + theme_bw(base_size = 12) + scale_color_manual(values = c("goldenrod", "darkred")) + theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white")) + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + ylab(TeX("standing height $\\hat{q}$ (eur-lat)")) + xlab("") +facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(threshold = c("strict" = "Strict Threshold", "loose" = "Liberal Threshold"))) + scale_shape_manual(values = c(8,16))
```

Sas-Amr - MCV
```{r}
pl3 <- df %>% filter(contrasts == "sas-amr", gtype == "LMM") %>% filter(phenotype == "MCV") %>% mutate(upper = q + 3.85, lower = q - 3.85) %>%  ggplot(aes(x = covar, y = q, color = covar, shape = sigP)) + geom_point(size = 3) +  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5) + theme_bw(base_size = 12) + scale_color_manual(values = c("goldenrod", "darkred")) + theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white")) + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + ylab(TeX("MCV $\\hat{q}$ (sas-amr)")) + xlab("") +facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(threshold = c("strict" = "Strict Threshold", "loose" = "Liberal Threshold"))) + scale_shape_manual(values = c(8,16))
```

Eng-Sct - SBP
```{r}
pl4 <- df %>% filter(contrasts == "England-Scotland", gtype == "LMM") %>% filter(phenotype == "SBP") %>% mutate(upper = q + 3.85, lower = q - 3.85) %>%  ggplot(aes(x = covar, y = q, color = covar, shape = sigP)) + geom_point(size = 3) +  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5) + theme_bw(base_size = 12) + scale_color_manual(values = c("goldenrod", "darkred")) + theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white")) + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + ylab(TeX("SBP $\\hat{q}$ (ENG-SCT)")) + xlab("") +facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(threshold = c("strict" = "Strict Threshold", "loose" = "Liberal Threshold"))) + scale_shape_manual(values = c(8,16))
```

RoI-Sct - SBP
```{r}
pl5 <- df %>% filter(contrasts == "RepublicOfIrelend-Scotland", gtype == "LMM") %>% filter(phenotype == "SBP") %>% mutate(upper = q + 3.85, lower = q - 3.85) %>%  ggplot(aes(x = covar, y = q, color = covar, shape = sigP)) + geom_point(size = 3) +  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5) + theme_bw(base_size = 12) + scale_color_manual(values = c("goldenrod", "darkred")) + theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white")) + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + ylab(TeX("SBP $\\hat{q}$ (RoI-SCT)")) + xlab("") +facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(threshold = c("strict" = "Strict Threshold", "loose" = "Liberal Threshold"))) + scale_shape_manual(values = c(8,16))
```

Sct-Wales - SBP
```{r}
pl6 <- df %>% filter(contrasts == "Scotland-Wales", gtype == "LMM") %>% filter(phenotype == "SBP") %>% mutate(upper = q + 3.85, lower = q - 3.85) %>%  ggplot(aes(x = covar, y = q, color = covar, shape = sigP)) + geom_point(size = 3) +  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5) + theme_bw(base_size = 12) + scale_color_manual(values = c("goldenrod", "darkred")) + theme(legend.position = "none", strip.background=element_rect(colour="black",fill="white")) + geom_hline(yintercept = 0, color = "red", linetype = "dashed") + ylab(TeX("SBP $\\hat{q}$ (SCT-WAL)")) + xlab("") +facet_grid(rows = NULL, cols = vars(gwas, threshold), labeller = labeller(threshold = c("strict" = "Strict Threshold", "loose" = "Liberal Threshold"))) + scale_shape_manual(values = c(8,16))
```

### Final Significant Results Figure 

```{r}
library(gridExtra)
library(grid)

# Define label grobs
label_a <- textGrob("a", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_b <- textGrob("b", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_c <- textGrob("c", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_d <- textGrob("d", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_e <- textGrob("e", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_f <- textGrob("f", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))
label_g <- textGrob("g", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontface = "bold", fontsize = 14))

# Wrap plots with labels
row_a <- arrangeGrob(plManhattan, top = label_a)
row_b <- arrangeGrob(pl1, top = label_b)
row_c <- arrangeGrob(pl2, top = label_c)
row_d <- arrangeGrob(pl3, top = label_d)
row_e <- arrangeGrob(pl4, top = label_e)
row_f <- arrangeGrob(pl5, top = label_f)
row_g <- arrangeGrob(pl6, top = label_g)

# Final arrangement: top_row + c, d, e stacked below
final_plot <- grid.arrange(row_a, row_b, row_c, row_d, row_e, row_f, row_g, 
                           ncol = 1,
                           heights = c(0.75, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5))

# Save
ggsave("~/Desktop/SigFig.png", final_plot, height = 18, width = 12)
```



## CoB in WBS 

```{r}
df <- fread("../plots/pga_test_resampled/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(type = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```

### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```


## CoB in ALL


```{r}
df <- fread("../plots/pga_test/InUKBB/ALL-R/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_InUKBB.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_InUKBB.txt") %>% mutate(type = "CoB") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Strict Threshold")
```


Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```


### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```



Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```


## HGDP in WBS


```{r}
df <- fread("../plots/pga_test/HGDP1KG/WBS/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(type = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```


Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```


### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```



Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```

## HGDP in ALL 

```{r}
df <- fread("../plots/pga_test/HGDP1KG/ALL/q_results.txt")
tmp2 <- fread("../plots/overlap_stats/H_HGDP1KG.txt")
tmp3 <- fread("../plots/error/common_rare_R2EO_HGDP1KG.txt") %>% mutate(type = "HGDP1KG") %>% filter(PC == 80) %>% select(contrasts, Ratio, gwas)

df <- left_join(df, tmp2) %>% mutate(adjuP = p.adjust(pval)) %>% mutate(sigP = case_when(adjuP < 0.05  ~ TRUE, adjuP >= 0.05 ~ F))
df <- left_join(df, tmp3)
```


### Q^2 vs H

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + scale_x_log10() + ggtitle("Loose Threshold")
```


### Q^2 vs Number of SNPs 

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = nsnp, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold")
```

### Q^2 vs H'

Strict threshold
```{r}
df %>% filter(threshold == "strict") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Strict Threshold") + xlab("H'")
```

Loose threshold
```{r}
df %>% filter(threshold == "loose") %>% ggplot(aes(x = (1-Ratio) * H, y = q^2, color = sigP)) + geom_point() + facet_grid(vars(covar), vars(gtype)) + ggtitle("Loose Threshold") + xlab("H'")
```

```{r}
df <- fread("~/strat2/plots/error/R2EO_InUKBB.txt")

df %>% filter(PC == 40) %>% ggplot(aes(x = H, y = Ratio, color = gwas)) + geom_point() + scale_x_log10() + theme_bw() + theme( strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab(TeX("proportion of variance explained by contrast in GWAS panel ($\\hat{H}$)")) 
```


```{r}
df <- fread("~/strat2/plots/error/500_R2EO_InUKBB.txt")


df %>% filter(PC == 40) %>% ggplot(aes(x = H, y = Ratio, color = gwas)) + geom_point() + scale_x_log10() + theme_bw() + theme( strip.background=element_rect(colour="black",fill="white"), axis.text.x = element_text(size = 10)) + xlab(TeX("proportion of variance explained by contrast in GWAS panel ($\\hat{H}$)"))
```
```

